It has been quite a journey at Stanford.
In one word, it was just really {\em fun} -- it was fun to think about, talk about, and write about the problems I loved working on. 
As I am putting a proper end to this journey, I am indebted to a lot of people for their help, support, advice, and memories.

To the Kwanjeong Educational Foundation and its staffs, for supporting my graduate studies financially -- it helped me concentrate on my research.

To my advisors, Yoav Shoham, Virginia Williams, and Ashish Goel, for their guidance and advice. From Yoav, more than anything, I learned how to be proactive -- ``ask for forgiveness, not permission'' is the saying that has changed my perspectives and habits. I thank Virginia and Ashish, who are officially my co-advisors, for agreeing to advise me after Yoav had decided to retire. Thanks to their advice and support, I was able to overcome what seemed to be the end of the road at the time. I had a few ideas about a few problems, but my ideas did not pan out exactly the way I wanted, and I was not sure how to proceed further. With Ashish, I focused more on application-driven problems with practical use in mind, and with Virginia, I dove into a deeper level of computational complexity than what I knew about it before I met her. As Yoav predicted, I got to experience the best of both worlds, and I really enjoyed it. So, again, I thank all three of my advisors for sharing their wisdom and guiding me through this journey.

To June, Junghwa, Soah, Yoonho, Jangryul, Minjung, Jaehyun, Fiona, Brian, Heeyoon, Jaehwan, Sean, Jihoon, Albert, Ernest, and Minyong, for a lot for great memories and fun times at Stanford. Special thanks to Albert, Ernest, and Minyong for agreeing to take part in my pop-up seminar series, which I thought of at Minyong's studio when we were not sober enough to talk about math-y stuff  -- thep pop-up seminars were informal but very inspiring. 

To Adrian Marple, Jacob Bank, and Eric Huang -- it was really fun to work with you guys. Hope we go to the all-you-can-eat-sushi place one more time.

To Sehyuk Park and Ikkjin Ahn, for feeding me interesting real-world problems they deal with at Moloco -- it was an uplifting digression from my day-to-day doctoral research.

To Bobby Kleinberg and Johannes Gehrke, for having me as a research assistant at Cornell University -- working with you two convinced me to pursue a doctoral degree, and I have no regret about it. Serendipity is the word I would use here -- it never occurred to me that I would pursue a doctoral degree (I was not even sure of the meaning of a doctoral degree at the time), but I realized that it was something I would love doing, all thanks to you. 

To Anand Bhaskar and SueYeon Chung, for motivating me to consider a doctoral degree in the first place. I will never forget the time when we stayed up all night at the CS Undergrad Lab in Upson Hall when we participated in the Mathematical Contest of Modeling -- I learned a lot from you, and I really enjoyed our time together, taking the challenges.

To my in-laws, Albert, MC, and WK, for their kindness, joy, and love.

To my parents, JL and MM, for their tremendous love and support. My ride began when my parents let me learn programming back in year 2000 even after I failed on the most basic test about computers. I loved it, and I have been loving it since -- I happened to have found my thing early on, thanks to my parents. 

To my daughter, Maya, for making my journey more exciting -- truly blessing. The first few months was really tough -- I suffered from paternal postnatal depression. It was a rough time for me as I did not want to meet with anyone outside home and was constantly worried about everything. It was slowly building up as days passed by when all I did the whole day was change of diapers, feedings, and what-not. I was not aware of it, or perhaps I was suppressing the feelings at the time, but now I know what it was. The most helpful thing for me, in retrospect, was the endless calls and texts from friends and colleagues, asking me to get out for lunch or coffee or whatever. I really thank you all for pulling me out of the abysmal void of my mind.

Last but never the least, to my wife, Alice, for her wisdom, love, and support. I will never forget how we spent our Spring Breaks on playing video games, watching Breaking Bad, and living in Minecraft. Everything becomes more fun to do with you.

Again, I thank all of you for making my journey a special and memorable one, and I surely enjoyed my ride. Now I am off to a new start!


\vspace{4 em}

원하에게: 

너를 ``기억할 장소가 존재하지 않기를 바란다''라고 한들 너를 아끼던 모두가 더 열심히 기억하고 추억할 것이다. 네 의지대로 떠났지만, 남겨진 우리는 우리의 의지대로 너를 기억할 것 같다. 다시 만 날 때까지 그 곳에서 즐기고 있길.
